\chapter{Introduction}

Advances in manufacturing and hardware design have made mobile robots more accessible for not only research and industrial purposes, but to the general public. However, autonomous robots have largely been limited to indoor, carefully controlled environments. Before autonomous robots can be more widely deployed, they must be able to accurately observe and represent unstructured, dynamic environments. Dense sensors such as light field cameras are becoming cheaper and lighter, promising to allow autonomous robots to acquire detailed measurements of these complex surroundings. To fully utilise the potential of these advancing technologies, improvements in observer design are required to generate more accurate and detailed descriptions of these environments from dense measurements.

One method of estimating the state of a complex environment is with an \textit{infinite-dimensional observer}. Typically, observers for infinite-dimensional systems are extensions of finite-dimensional Luenberger observers. Unfortunately, this design approach is only able to guarantee convergence for linear systems. Developing a theory of symmetry-preserving, infinite-dimensional observers would simplify the design process for nonlinear systems and result in observers with improved convergence properties.

This research project serves as an initial exploration into the design of symmetry-preserving, infinite-dimensional observers. An infinite-dimensional system will be simplified and an observer will be designed to estimate a finite-dimensional state. The potential for a symmetry-preserving, infinite-dimensional observer to improve performance will be explored. 

This report presents the novel implementation of an observer to estimate the state of a rigid cube from range measurements. Section \ref{sec:literature} reviews the current state of observer design methods for infinite-dimensional systems. Recent work in the development of design methodologies for symmetry preserving observers is described. Particular attention is paid to a dense optical flow estimator that will be particularly relevant to this research.
 
Chapter \ref{chap:background} provides theory on Lie groups, rigid body state representation and state observer concepts that will be applied in the observer design.

The cube state estimation problem that is the focus of this research is described in detail in Chapter \ref{chap:problem}. The place of this problem within the larger area of symmetry-preserving, infinite-dimensional observer research is defined.  

Chapter \ref{chap:simulation} provides a detailed description of the simulation implementation, including the observer update function design. The performance of the observer in estimating the state of stationary and moving cubes is assessed. It is shown that almost global convergence is achieved for orientation and size correction for stationary cubes, though the position update only converges in special cases.

Chapter \ref{chap:experiment} details steps taken to experimentally validate the results of the simulation. Range measurement data is collected and used to develop an error distribution model for the Hokuyo UBG-04LX-F01 sensor. Range measurements of a cube of known trajectory are taken for future performance testing.

\input{literature}
